---
title: "test"
author: "Jong-Hoon Kim"
date: "1/23/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
# theta <- data.table::fread("inst/extdata/theta.csv")
# y0 <- data.table::fread("inst/extdata/y0.csv")
# usethis::use_data(theta, overwrite = T)
# usethis::use_data(y0, overwrite = T)
# devtools::document()
# y1 = y0$val
# names(y1) = y0$name
# y0 = y1

```

# Check functions
## process_model
```{r}
nparticle = 100
tend = 20
# beta <- theta[name =="R0", val] * theta[name =="gamma", val]
beta = theta[["R0"]] * theta[["gamma"]]
# volatility for beta (random walk)
beta_vol <- matrix(rnorm(nparticle * tend, mean = 0, sd = theta[["betavol"]]), nrow = tend)
beta_vol[1,] <- exp(beta_vol[1,]) * beta # initial beta
y <- data.frame(matrix(0, nrow = nparticle, ncol = length(y0)))
names(y) <- names(y0)
y[,"S"] <- 1e7
y[,"I"] <- 100
yt <- y1 <- process_model(params = theta, y = y, beta = beta_vol[1,])
for (i in 1:100) {
  yt <- process_model(params = theta, y = yt, beta = beta_vol[1,])
}
head(y1)
head(yt)
head(beta_vol[1,])
```
## Rt data generation
```{r}
nt <- 200 # days
presymp_infect <- TRUE
param <- list()
param$sigma <- 1/4
param$gamma <- 1/4
param$R <- c(rep(1.2, nt/4), rep(1.4, nt/4), rep(0.8, nt/4), rep(1.2, nt/4))
param$presymp_infect <- FALSE
times <- 0:(nt-1) # simulation times
y0 <- c(S = 1e7, E1 = 40, E2 = 40, I = 100, R = 0, CE1 = 0,  CE2 = 0, CI = 0) # initial values

library(deSolve) 
library(tidyverse)

ode(func = se1e2ir, y = y0, times = times, parms = param) %>%
  as.data.frame() -> out

inf_daily <- diff(out$CE1)
onset_daily <- diff(out$CI)
confirm_daily <- diff(out$R)

df <- data.frame(t = 1:length(param$R), 
                 daily_R_true = param$R, 
                 daily_infect = c(0, inf_daily),
                 daily_onset = c(0, onset_daily),
                 daily_confirm = c(0, confirm_daily))

df %>% 
  pivot_longer(cols = -t, names_to = "var") %>%
  filter(var != "daily_R_true") %>% 
  ggplot(aes(t, value, color = var)) +
  geom_line() +
  labs(x = "time (day)", y = "number of individuals")
# Rt_data <- df
# usethis::use_data(Rt_data, overwrite = T)
```

## particle_filter - sample
```{r}
pf <- particle_filter(data_type = "infection", npart = 1e2)
pf$lik_overall_average
df <- cbind(Rt_data[,1:2], Rt_pf = pf$trace$beta/theta[["gamma"]])
suppressPackageStartupMessages(library(tidyverse))
df %>% pivot_longer(cols = -t) %>% 
  ggplot(aes(t, value, color = name)) + 
  geom_line()
```

## particle_filter - filtered distribution
```{r}
theta[["betavol"]] = 0.2
pf <- particle_filter(data_type = "symptom onset", npart = 1e3)
pf$lik_overall_average
pr <- c(0.025, 0.25, 0.5, 0.75, 0.975)
Rt_quantile <- as.data.frame(t(apply(pf$beta_filtered/theta[["gamma"]], 1, function(x) {quantile(x, pr)})))
Rt_mean <- data.frame(mean = rowMeans(pf$beta_filtered/theta[["gamma"]]))
df <- cbind(Rt_quantile, Rt_mean, Rt_data[,1:2])

suppressPackageStartupMessages(library(tidyverse))
ggplot(df, aes(x = t)) +
  geom_ribbon(aes(ymin = `2.5%`, ymax = `97.5%`), fill = "grey") +
  geom_line(aes(y = `50%`), size = 1) + 
  geom_line(aes(y = mean), color = "steelblue", size = 1) + 
  geom_point(aes(y = daily_R_true), color = "darkred") +
  labs(title = "R(t) via particle filtering", y = "R(t)", x = "Day")

```

## Particle filtering and smoothing
$$X(t_k) | Y(t_1)=y^*_1, \dots,\; Y(t_n)=y^*_n,\; X_k | Y_1=y_1^*, \dots,\; Y_n=y_n^*$$
```{r}
res <- run_model(data_type = "confirmation", rep = 1e2, npart = 1e3, systematic_resampling = T)
        
pr <- c(0.025, 0.25, 0.5, 0.75, 0.975)
Rt_quantile <- as.data.frame(t(apply(res$Rt, 1, function(x) {quantile(x, pr)})))
library(data.table)
Rt_mean <- data.frame(mean = rowMeans(res$Rt), 
                      rollmean = frollmean(Rt_quantile[,3], n=7))
df <- cbind(Rt_quantile, Rt_mean, Rt_data[,1:2])
library(tidyverse)
ggplot(df, aes(x = t)) +
  geom_ribbon(aes(ymin = `2.5%`, ymax = `97.5%`), fill = "grey") +
  geom_line(aes(y = `50%`), size = 1) + 
  geom_line(aes(y = mean), color = "steelblue", size = 1) + 
  geom_point(aes(y = daily_R_true), color = "darkred") +
  labs(title = "R(t) via particle filtering with smooothing", y = "R(t)", x = "Day")
```

## repeat particle_filter and sample (backward sampling)
```{r}
res <- run_model(data_type = "confirmation", rep = 1e3, npart = 1e3, 
                 systematic_resampling = F)
          
pr <- c(0.025, 0.25, 0.5, 0.75, 0.975)
Rt_quantile <- as.data.frame(t(apply(res$Rt, 1, function(x) {quantile(x, pr)})))
library(data.table)
Rt_mean <- data.frame(mean = rowMeans(res$Rt), 
                      rollmean = frollmean(Rt_quantile[,3], n=7))
df <- cbind(Rt_quantile, Rt_mean, Rt_data[,1:2])
suppressPackageStartupMessages(library(tidyverse))
ggplot(df, aes(x = t)) +
  geom_ribbon(aes(ymin = `2.5%`, ymax = `97.5%`), fill = "grey") +
  geom_line(aes(y = `50%`), size = 1) + 
  geom_line(aes(y = mean), color = "steelblue", size = 1) + 
  geom_point(aes(y = daily_R_true), color = "darkred") +
  labs(title = "R(t) via particle filtering with backward sampling", y = "R(t)", x = "Day")

# library(data.table)
# df <- data.frame(pf_Rt_rollmean = frollmean(Rt_quantile[,3], n=7))
# df <- cbind(Rt_data[,1:2], df)
# df %>% 
#   pivot_longer(cols = - t) %>%
#   ggplot(aes(t, value, color = name)) +
#   geom_line() +
#   geom_point()
```
## Maximum likelihood estimation test
```{r}
# devtools::load_all(".")
mle <- maxlik(data_type = "confirmation")

df <- cbind(Rt_data[,1:2], (mle$beta_trace/theta[name == "gamma", val]))
suppressPackageStartupMessages(library(tidyverse))
df %>% 
  pivot_longer(cols = - t) %>%
  ggplot(aes(t, value, color = name)) +
  geom_line() +
  geom_point() +
  labs(title = "R(t) via maximum likelihood", y = "R(t)", x = "Day")

```

## RcppArmadillo 3D array test 
```{r}
library(Rcpp)
library(RcppArmadillo)
sourceCpp("cpp/test.cpp")
a = array(1, dim=c(12,12,12))
s = R_C_sum(a)
# > (s = R_C_sum(a))
# [1] 1728
# > 12**3
# [1] 1728
a2 = C_R_times_two(a)
# > class(a2)
# [1] "array"
# > dim(a2)
# [1] 12 12 12
# > a2[1:2,1:2,1:2]
# , , 1
# 
#      [,1] [,2]
# [1,]    2    2
# [2,]    2    2
# 
# , , 2
# 
#      [,1] [,2]
# [1,]    2    2
# [2,]    2    2

arr = array(0, dim=c(2,2,2))
cube_with_ones(arr)
```

## Rcpp integer casting test
```{r}
library(Rcpp)
library(RcppArmadillo)
sourceCpp("cpp/test.cpp")
count_dt(tbegin = 10, tend = 12, dt = 0.1)
```

## Rcpp process model test
```{r}
library(Rcpp)
library(RcppArmadillo)
sourceCpp("cpp/process_model_cpp.cpp")
y0mat = matrix(y0, nrow = 1)
res = process_model_cpp(params = theta, y = y0mat, tbegin = 0, tend = 2, dt = 0.1, beta = 0.3)

beta = 0.5
tend = 200
res = matrix(NA, nrow = tend + 1, ncol = length(y0))
m = matrix(y0, nrow = 1)
res[1, ] = m[1, ]
for (i in 1:tend){
  m = process_model_cpp(params = theta, y = m, tbegin = 0, tend = 1, dt = 0.2, beta = beta)
  res[i + 1, ] = m[1, ]
}
res = as.data.frame(res)
names(res) = c("S", "E1", "E2", "I", "R", "CE1", "CE2", "CI")
suppressMessages(library(tidyverse))
res %>% mutate(day = (1:n() - 1)) %>% 
  pivot_longer(cols = - day) %>% 
  dplyr::filter(name == "I" | name == "R") %>% 
  ggplot(aes(x = day)) +
  geom_line(aes(y = value, color = name))


## outbreak size comparision - to have a sense that numerical solutions are reasonable
R0 = beta / theta[["gamma"]]
outbreak_size <- function (R0, final) {
   return (1 - exp( - R0 * final) - final)
}
sols <- rootSolve::multiroot( outbreak_size, c(0, 1), R0=2 )
max(sols$root) # 0 is also a solution
y0[["S"]] * max(sols$root)
res[tend+1, "R"]
```

## Testing particle filter implemented in C++ 
## particle_filter_cpp - filtered distribution
```{r}
library(Rcpp)
library(RcppArmadillo)
sourceCpp("cpp/particle_filter_cpp.cpp")
devtools::load_all(".")
theta[["betavol"]] = 0.2
dmat = data.matrix(Rt_data)
pf = particle_filter_cpp(theta, y0, dmat, "infection", 1e3L, 200L, 0.1)

pf$lik_overall_average
pr <- c(0.025, 0.25, 0.5, 0.75, 0.975)
Rt_quantile <- as.data.frame(t(apply(pf$beta_filtered/theta[["gamma"]], 1, function(x) {quantile(x, pr)})))
Rt_mean <- data.frame(mean = rowMeans(pf$beta_filtered/theta[["gamma"]]))
df <- cbind(Rt_quantile, Rt_mean, Rt_data[,1:2])

suppressPackageStartupMessages(library(tidyverse))
ggplot(df, aes(x = t)) +
  geom_ribbon(aes(ymin = `2.5%`, ymax = `97.5%`), fill = "grey") +
  geom_line(aes(y = `50%`), size = 1) + 
  geom_line(aes(y = mean), color = "steelblue", size = 1) + 
  geom_point(aes(y = daily_R_true), color = "darkred") +
  labs(title = "R(t) via particle filtering", y = "R(t)", x = "Day")
```


## repeat particle_filter and sample (backward sampling)
```{r}
library(Rcpp)
library(RcppArmadillo)
sourceCpp("cpp/particle_filter_cpp2.cpp")
devtools::load_all(".")
theta[["betavol"]] = 0.2

res <- run_model_cpp(data_type = "infection", rep = 1e2, npart = 1e3)
          
pr <- c(0.025, 0.25, 0.5, 0.75, 0.975)
Rt_quantile <- as.data.frame(t(apply(res$Rt, 1, function(x) {quantile(x, pr)})))
library(data.table)
Rt_mean <- data.frame(mean = rowMeans(res$Rt), 
                      rollmean = frollmean(Rt_quantile[,3], n=7))
df <- cbind(Rt_quantile, Rt_mean, Rt_data[,1:2])
suppressPackageStartupMessages(library(tidyverse))
ggplot(df, aes(x = t)) +
  geom_ribbon(aes(ymin = `2.5%`, ymax = `97.5%`), fill = "grey") +
  geom_line(aes(y = `50%`), size = 1) + 
  geom_line(aes(y = mean), color = "steelblue", size = 1) + 
  geom_point(aes(y = daily_R_true), color = "darkred") +
  labs(title = "R(t) via particle filtering with backward sampling", y = "R(t)", x = "Day")
```
## performance comparison
```{r}
library(Rcpp)
library(RcppArmadillo)
sourceCpp("cpp/particle_filter_cpp.cpp")
devtools::load_all(".")
library(microbenchmark)
microbenchmark(run_model_cpp(data_type = "infection", rep = 1, npart = 1e3, dt = 0.1), 
               run_model(data_type = "infection", rep = 1, npart = 1e3, dt = 0.1), times = 50L)

```

## Parallel 
```{r}
library(parallel)
library(doParallel)
library(foreach)
ncores = detectCores()
cl <- makeCluster(getOption("cl.cores", ncores-2))
doParallel::registerDoParallel(cl)
system.time({
r <- foreach (i=1:1000, .combine = cbind, .packages ="particlefilter", .inorder = F) %dopar% {
  run_model(params = theta, y = y0, data = Rt_data, data_type = "infection",
            rep = 1, npart = 1e3, tend = 200, dt = 0.1, 
            systematic_resampling = FALSE)
}})
parallel::stopCluster(cl)

rt = r[9,] # Rt
rtdf = data.frame(matrix(unlist(rt), ncol=length(rt), byrow=F))
pr <- c(0.025, 0.25, 0.5, 0.75, 0.975)
Rt_quantile <- as.data.frame(t(apply(rtdf, 1, function(x) {quantile(x, pr)})))
library(data.table)
Rt_mean <- data.frame(mean = rowMeans(rtdf), 
                      rollmean = frollmean(Rt_quantile[,3], n=7))
df <- cbind(Rt_quantile, Rt_mean, Rt_data[,1:2])
suppressPackageStartupMessages(library(tidyverse))
ggplot(df, aes(x = t)) +
  geom_ribbon(aes(ymin = `2.5%`, ymax = `97.5%`), fill = "grey") +
  geom_line(aes(y = `50%`), size = 1) + 
  geom_line(aes(y = mean), color = "steelblue", size = 1) + 
  geom_point(aes(y = daily_R_true), color = "darkred") +
  labs(title = "R(t) via particle filtering with backward sampling", y = "R(t)", x = "Day")

# library(tidyverse)
# rtdf %>% 
#   mutate(day = 1:n() - 1) %>% 
#   pivot_longer(cols = - day) %>%
#   ggplot(aes(x = day))+
#   geom_line(aes(y = value))
  
  

  
# cl <- makeCluster(getOption("cl.cores", 10))
# doParallel::registerDoParallel(cl)
# clusterExport(cl = cl, varlist = c("run_model_cpp", "y0", "theta", "Rt_data", "particle_filter_cpp"))
# # system.time({
# # r <- foreach (i=1:50, .combine=cbind, .packages = "particlefilter") %dopar% {
# #   run_model(data_type = "infection", rep = 1, npart = 1e3, dt = 0.1)
# # }})
# system.time({
# r <- foreach (i=1:50, .combine=cbind) %dopar% {
#   run_model_cpp(data_type = "infection", rep = 1, npart = 1e3, dt = 0.1)
# }})
# parallel::stopCluster(cl)

```

## KDCA data
### Data cleaning
```{r}
library(readxl)
suppressMessages(library(tidyverse))
d <- read_xlsx("data/covid_kdca.xlsx")
names(d) = c("id", "nationality", "sex", "age", "sido", "sigungu", "date_symptom_onset", "date_diagnosis", "date_report", "route_infection", "country_orgin", "id_infector", "occupation")
d1 <- d %>% filter(route_infection == "국내")%>% 
  select(date_diagnosis) %>% 
  group_by(date = date_diagnosis) %>% 
  summarise(daily_confirm = n())
d2 <- data.frame(date = seq(as.Date("2020-01-03"), as.Date("2021-02-23"), by = "day"))
d2$t <- 1:nrow(d2)
d3 <- left_join(d2, d1, by = "date")
dat <- d3 %>% mutate(daily_confirm = ifelse(is.na(daily_confirm), 0, daily_confirm))
## name standardization
ggplot(dat, aes(t, daily_confirm)) +
  geom_line()

```

### Data fitting
```{r}
theta[["betavol"]] = 0.6
ynew <- c(S = 5*1e7, E1 = 40, E2 = 40, I = 1e2, R = 0, CE1 = 0, CE2 = 0, CI = 0)  
pf <- particle_filter(ydata = dat, data_type = "confirmation", npart = 1e4, tend = 418)

pf$lik_overall_average
pr <- c(0.025, 0.25, 0.5, 0.75, 0.975)
Rt_quantile <- as.data.frame(t(apply(pf$beta_filtered/theta[["gamma"]], 1, function(x) {quantile(x, pr)})))
Rt_mean <- data.frame(mean = rowMeans(pf$beta_filtered/theta[["gamma"]]))
df <- cbind(Rt_quantile, Rt_mean, dat[,c("t", "date")])

suppressMessages(library(tidyverse))
ggplot(df, aes(x = t)) +
  geom_ribbon(aes(ymin = `2.5%`, ymax = `97.5%`), fill = "grey") +
  geom_line(aes(y = `50%`), size = 1) + 
  geom_line(aes(y = mean), color = "steelblue", size = 1) + 
  labs(title = "R(t) via particle filtering", y = "R(t)", x = "Day") +
  geom_hline(yintercept = 1, color = "darkred", linetype = "dotted") + 
  scale_y_continuous(limits = c(0, 100))

ci <- t(pf$latent_var_filtered[,,8])
daily_confirm <- diff(ci)
for (i in 1:ncol(daily_confirm)){ 
  if (sum(daily_confirm[, i] < 0) > 0) {
    cat("part =", i, "\n"); stop("negative")
  }
}
plot(daily_confirm[,1], cex = 0.4)
abline(h = 0, col = 2)
points(dat$t, dat$daily_confirm, col = 3, cex = 0.5)

sum(daily_confirm < 0)
ci <- t(apply(pf$latent_var_filtered[,,8], 2, function(x) {quantile(x, pr)}))

daily_confirm <- diff(ci)
```

### Data fitting
```{r}
library(parallel)
library(doParallel)
library(foreach)
ncores = detectCores()
cl <- makeCluster(getOption("cl.cores", ncores-2))
doParallel::registerDoParallel(cl)
clusterExport(cl = cl, varlist = c("dat"), envir = .GlobalEnv)
system.time({
r <- foreach (i=1:1000, .combine = cbind, .packages ="particlefilter", .inorder = F) %dopar% {
  run_model(params = theta, y = y0, data = dat, data_type = "confirmation",
            rep = 1, npart = 1e3, tend = 200, dt = 0.1, 
            systematic_resampling = FALSE)
}})
parallel::stopCluster(cl)

```

## Testing arma::cube can be access through _ just like in Rcpp::NumericMatrix
```{r}
library(Rcpp)
library(RcppArmadillo)
sourceCpp("cpp/test.cpp")
# a = array(1:12, c(2,3,2))
# a[, , 1]
# (a = export_array(Q = a, slice = 0))
# # (m = export_mat(Q = a, slice = 0))
# (v = export_vec(Q = a, col = 0, slice = 0))
# (i = to_int(x = c(1.0, 2.2)))
# m = matrix(1:12, nrow = 3)
# m[,1]
# (v = export_vec_from_mat(m = m, col = 0))
# (ch = add_x("mr. "))
# all_x("x")
# all_x("y")
int_sample(10)
int_wt_sample(10, wt = c(0.8,0.2,rep(0,8)))
sample_test(10, wt = c(0.8,0.2,rep(0,8)))
```

## testing assign_weights function
```{r}
library(Rcpp)
library(RcppArmadillo)
sourceCpp("cpp/assign_weights_cpp.cpp")
st_now = data.matrix(data.frame(S=rep(1,10), E1 = rep(1,10), E2 = rep(1,10), I = rep(1,10), R = 1:10, CE1 = 101:110, CI = 101:110))
st_before = data.matrix(data.frame(S=rep(1,10), E1 = rep(1,10), E2=rep(1,10), I = rep(1,10), R = 1:10-1, CE1 = round(seq(1, 100, length.out = 10)), CI = round(seq(1, 100, length.out = 10))))

wt = assign_weights_cpp(st_now, st_before, 53L, data.matrix(Rt_data), "infection")
all.equal(wt[1], dpois(101,100))
```

## testing particle_filter_cpp function
```{r}
library(Rcpp)
library(RcppArmadillo)
sourceCpp("cpp/particle_filter_cpp.cpp")
```

##Conditional log likelihood
The estimated conditional log likelihood from a fitted model.
The conditional likelihood is defined to be the value of the density
$$ Y(t_k) | Y(t_1),\dots,Y(t_{k-1}) Y_k | Y_1,\dots,Y_{k-1}$$ 
evaluated at $$Y(t_k) = y^*_k Y_k = y_k^*$$.

Here, \eqn{Y(t_k)}{Yk} is the observable process, and \eqn{y^*_k}{yk*} the data, at time \eqn{t_k}.

Thus the conditional log likelihood at time \eqn{t_k} is
\deqn{\ell_k(\theta) = \log f[Y(t_k)=y^*_k \vert Y(t_1)=y^*_1, \dots, Y(t_{k-1})=y^*_{k-1}],}{ell_k(theta)=log f[Yk = yk* | Y1=y1*, \dots, Y(k-1)=y(k-1)*],}
where \eqn{f} is the probability density above.

## pomp package
```{r}
library(tidyverse)
read_csv(paste0("https://kingaa.github.io/sbied/stochsim/", "Measles_Consett_1948.csv")) %>%
select(week,reports=cases) -> meas
meas %>% as.data.frame() %>% head()

sir_step <- function (S, I, R, N, Beta, mu_IR, delta.t, ...) {
  dN_SI <- rbinom(n=1,size=S,prob=1-exp(-Beta*I/N*delta.t))
  dN_IR <- rbinom(n=1,size=I,prob=1-exp(-mu_IR*delta.t))
  S <- S - dN_SI
  I <- I + dN_SI - dN_IR
  R <- R + dN_IR
  c(S = S, I = I, R = R)
}

sir_rinit <- function (N, eta, ...) {
  c(S = round(N*eta), I = 1, R = round(N*(1-eta)))
}

library(pomp)
meas %>%
pomp(times="week",t0=0,
rprocess=euler(sir_step,delta.t=1/7),
rinit=sir_rinit
) -> measSIR

sir_dmeas <- function (reports, H, rho, log, ...) {
  dbinom(x=reports, size=H, prob=rho, log=log)
}
sir_rmeas <- function (H, rho, ...) {
  c(reports=rbinom(n=1, size=H, prob=rho))
}
measSIR %>%
  pomp(
  rmeasure=sir_rmeas,
  dmeasure=sir_dmeas
) -> measSIR

measSIR %>%
  simulate(
  params=c(Beta=7.5,mu_IR=0.5,rho=0.5,eta=0.03,N=38000),
  nsim=20,format="data.frame",include.data=TRUE
) -> sims
sims %>%
  ggplot(aes(x=week,y=reports,group=.id,color=.id=="data"))+
  geom_line()+
  guides(color=FALSE)
```

```{r}
library(pomp)
library(tidyverse)
seir_step <- Csnippet("
  double N = S + E1 + E2 + I + R;
  double dN_SE1 = beta * S * I / N * dt;
  double dN_E1E2 = 2 * sigma * E1 * dt;
  double dN_E2I = 2 * sigma * E2 * dt;
  double dN_IR = gamma * I * dt;
  S +=  - dN_SE1;
  E1 += dN_SE1 - dN_E1E2;
  E2 += dN_E1E2 - dN_E2I;
  I += dN_E2I - dN_IR;
  R += dN_IR;
  H += dN_SE1;
")

seir_rinit <- Csnippet("
  S = S_0;
  E1 = E1_0;
  E2 = E2_0;
  I = I_0;
  R = R_0;
  H = H_0;
")

seir_dmeas <- Csnippet("
  lik = dpois(daily_infect, H, give_log);
")

seir_rmeas <- Csnippet("
  daily_infect = rpois(H);
")

dat <- data.frame(day = Rt_data$t, daily_infect = Rt_data$daily_infect)

dat %>% pomp(
  times = "day", 
  t0 = 1,
  rprocess = euler(seir_step, delta.t = 0.2),
  rinit = seir_rinit,
  rmeasure = seir_rmeas,
  dmeasure = seir_dmeas,
  accumvars = "H",
  statenames = c("S", "E1", "E2", "I", "R", "H"),
  paramnames = c("beta", "sigma", "gamma", "S_0", "E1_0", "E2_0", "I_0", "R_0", "H_0")) -> covid_seir

covid_seir %>%
  simulate(params = c(beta=1.2*0.25, sigma=0.25, gamma=0.25,
                      S_0=1e7, E1_0=40, E2_0=40, I_0=100, R_0=0, H_0=0),
nsim = 20, format="data.frame", include.data=TRUE) -> sims

sims %>%
  ggplot(aes(x=day,y=daily_infect, group=.id, color=.id=="data"))+
  geom_line()+
  guides(color=FALSE)

covid_seir %>%
  pfilter(Np=1000, paramnames="beta") -> pfrick

```

```{r}
library(pomp)
library(tidyverse)
seir_step <- Csnippet("
  double N = S + E1 + E2 + I + R;
  double dN_SE1 = beta * S * I / N * dt;
  double dN_E1E2 = 2 * 0.25 * E1 * dt;
  double dN_E2I = 2 * 0.25 * E2 * dt;
  double dN_IR = 0.25 * I * dt;
  S +=  - dN_SE1;
  E1 += dN_SE1 - dN_E1E2;
  E2 += dN_E1E2 - dN_E2I;
  I += dN_E2I - dN_IR;
  R += dN_IR;
  H += dN_SE1;
")

seir_rinit <- Csnippet("
  S = 1e7;
  E1 = 40;
  E2 = 40;
  I = 100;
  R = 0;
  H = 0;
")

seir_dmeas <- Csnippet("
  lik = dpois(daily_infect, H, give_log);
")

seir_rmeas <- Csnippet("
  daily_infect = rpois(H);
")

dat <- data.frame(day = Rt_data$t, daily_infect = as.integer(Rt_data$daily_infect))

dat %>% pomp(
  times = "day", 
  t0 = 1,
  rprocess = euler(seir_step, delta.t = 0.2),
  rinit = seir_rinit,
  rmeasure = seir_rmeas,
  dmeasure = seir_dmeas,
  accumvars = "H",
  statenames = c("S", "E1", "E2", "I", "R", "H"),
  paramnames = c("beta")) -> covid_seir

covid_seir %>%
  simulate(params = c(beta=1.2*0.25),
nsim = 20, format="data.frame", include.data=TRUE) -> sims

sims %>%
  ggplot(aes(x=day,y=daily_infect, group=.id, color=.id=="data"))+
  geom_line()+
  guides(color=FALSE)

pf = pfilter(covid_seir, Np=1000, params = c(beta = 0.3), 
             filter.mean = T, filter.traj = T, verbose = T) 
pf@filter.mean %>% 
  t() %>%
  as.data.frame() %>%
  mutate(time = 1:n()) %>% 
  pivot_longer(-time) %>%
  filter(name == "I") %>% 
  ggplot(aes(time, value)) +
  geom_line(aes(color = name))
  
```      

```{r}
ewmeas %>%
  subset(time < 1952) %>%
  pomp(
    times="time",t0=1948,
    rprocess=euler(
      Csnippet("
        int nrate = 6;
        double rate[nrate];	// transition rates
        double trans[nrate];	// transition numbers
        double dW;

        // gamma noise, mean=dt, variance=(sigma^2 dt)
        dW = rgammawn(sigma,dt);

        // compute the transition rates
        rate[0] = mu*pop;	// birth into susceptible class
        rate[1] = (iota+Beta*I*dW/dt)/pop; // force of infection
        rate[2] = mu;		// death from susceptible class
        rate[3] = gamma;	// recovery
        rate[4] = mu;		// death from infectious class
        rate[5] = mu; 		// death from recovered class

        // compute the transition numbers
        trans[0] = rpois(rate[0]*dt);	// births are Poisson
        reulermultinom(2,S,&rate[1],dt,&trans[1]);
        reulermultinom(2,I,&rate[3],dt,&trans[3]);
        reulermultinom(1,R,&rate[5],dt,&trans[5]);

        // balance the equations
        S += trans[0]-trans[1]-trans[2];
        I += trans[1]-trans[3]-trans[4];
        R += trans[3]-trans[5];
      "),
      delta.t=1/52/20
    ),
    rinit=Csnippet("
      double m = pop/(S_0+I_0+R_0);
      S = nearbyint(m*S_0);
      I = nearbyint(m*I_0);
      R = nearbyint(m*R_0);
    "),
    paramnames=c("mu","pop","iota","gamma","Beta","sigma",
      "S_0","I_0","R_0"),
    statenames=c("S","I","R"),
    params=c(mu=1/50,iota=10,pop=50e6,gamma=26,Beta=400,sigma=0.1,
      S_0=0.07,I_0=0.001,R_0=0.93)
  ) -> ew1
```

